---
title: "Mt Car Dataset Analysis"
author: "Manik Tomar & 24215275"
format: 
  pdf:
    pdf-engine: xelatex
    toc: true
    number-sections: true
editor: visual
execute:
  echo: false
  error: false
---

\newpage

# Introduction

The **`mtcars`** dataset comes built-in with R and contains fuel consumption and design specifications for 32 cars. It includes variables such as **miles per gallon (*mpg*), number of cylinders (*cyl*), horsepower (*hp*), weight (*wt*), and more.**

Some key variables in this dataset include:

-   **mpg**: Miles/(US) gallon (fuel efficiency)
-   **hp**: Gross horsepower (engine power)
-   **wt**: Weight of the car (in 1000 lbs)

# Summary Statistics

Below is a summary of key variables — *mpg*, *hp*, and *wt* — presented in a neat table using `knitr::kable()`.This gives us more idea about the dataset and then we can perform further Analysis.

```{r}
# Load necessary library
library(knitr)

# Summary statistics
summary_df <- summary(mtcars[, c("mpg", "hp", "wt")])
kable(summary_df, caption = "Summary Statistics of mpg, hp, and wt in mtcars")

```

## From the summary:

-   **Fuel efficiency (mpg)** ranges from **10.4 to 33.9**, with a mean of around **20 mpg**. This indicates a wide variation in car mileage among different models.
-   **Horsepower (hp)** varies significantly, from as low as **52** to as high as **335**, suggesting some cars are much more powerful than others.
-   The **weight (wt)** of the cars ranges from about **1.5 to 5.4 (in 1000 lbs)**, which can also affect fuel efficiency and power.

***Since there is a variation in the fuel effiency(mpg) and Horsepower lets find out how it is related to each other***

# MPG vs Horsepower

```{r}
# Load library
library(ggplot2)

# Create scatterplot
ggplot(mtcars, aes(x = hp, y = mpg)) +
  geom_point(color = "orange", size = 2) +
  labs(
    title = "Fuel Efficiency vs Horsepower",
    x = "Horsepower (hp)",
    y = "Miles per Gallon (mpg)"
  ) +
  theme_minimal()

```

-   The scatterplot shows a ***negative relationship*** between ***horsepower and fuel efficiency***:

    -   Cars with ***higher horsepower tend to have lower mpg***.

    -   This is ***expected as powerful engines consume more fuel***.

    -   The pattern suggests that there is a ***trade-off between performance (hp) and fuel economy (mpg)***.
